Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-05-30 SDSS Presentation/SDSS Presentation.Rmd
Updates for 5.30.2022

---
title: "Exploring Rural Shrink Smart Through Guided Discovery Dashboards"
author: "Denise R. Bradford and Susan R. Vanderplas, PhD"
date: "June 8, 2022"
output:
  xaringan::moon_reader:
    seal: false
    includes:
      after_body:
        "js-addins.html"
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default", "metropolis-fonts", "metropolis" ,"css/modal.css", "css/sizeformat.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightlines: true
      countIncrementalSlides: true
---
class:title-slide-custom

@@

<br><br>
# Exploring Rural Shrink Smart Through Guided Discovery Dashboards
## SDSS 2022
### June 8, 2022
#### Denise Bradford
#### Department of Statistics, University of Nebraska - Lincoln 
##### `r fa("envelope", fill = "black")` [denise.bradford@huskers.unl.edu](denise.bradford@huskers.unl.edu) 
##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/) 
##### `r fa("github", fill = "black")` [drbradford12](https://github.com/drbradford12)

???
Thank you for inviting me for a visit today and for attending my talk.

@@

+ Introduction to Rural Shrink Smart Project with Iowa State

@@

## Rural Smart Shrinkage

<!-- Insert the shrink smart Logo -->

@@

--

@@

---
## Complications

@@

---
## Complications

@@

---
## Basic Idea

@@

- Assess usability, intelligibility and adaptability to new statistical methodology.

- .emph.cerulean[Exploratory Data Analysis (EDA)]: 
    - Incorporate this information into an exploratory dashboard
    - Change up design of the dashboard to manage the easy of adaptability across our rural towns

---
## Guiding Dashboard Design Principles

- The town leaders are the focus audience; thus, the town itself should be the central focus of the app.
- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions to offering services that enhance resident quality of life.
- Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.
- Allow for more detailed data and feature requests to improve the dashboard design over time.

--

We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration

---
## Statistical Analysis

@@

- Supervised analysis
    - Which variables are most associated with subjectively reported quality of life (in the 99 towns in the ISTP)?

---
## Visualization

@@

>>>>

---
title: "Exploring Rural Shrink Smart Through Guided Discovery Dashboards"
author: "Denise R. Bradford and Susan R. Vanderplas, PhD"
date: "June 8, 2022"
output:
  xaringan::moon_reader:
    seal: false
    includes:
      in_footer: footer.html
      after_body:
        "js-addins.html"
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["default", "metropolis-fonts", "metropolis" ,"css/modal.css", "css/sizeformat.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightlines: true
      countIncrementalSlides: true
---
class:title-slide-custom

@@

<br><br>
# Exploring Rural Shrink Smart Through Guided Discovery Dashboards
## SDSS 2022
### June 8, 2022
#### Denise Bradford
#### Department of Statistics, University of Nebraska - Lincoln 
##### `r fa("envelope", fill = "black")` [denise.bradford@huskers.unl.edu](denise.bradford@huskers.unl.edu) 
<!-- ##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/)  -->
##### `r fa("github", fill = "black")` [drbradford12](https://github.com/drbradford12)

<br><br><br>
.medium.emph[
Theme designed by Emily A. Robinson [www.emilyarobinson.com](https://www.emilyarobinson.com/)
]
???
Thank you for inviting me for a visit today and for attending my talk.

@@

+ Introduction to Rural Shrink Smart Project with Iowa State University

@@

# Rural Shrink Smart Project
.center[
![](Shrink_Smart_logo.png)
]
.center[
(https://ruralshrinksmart.org/)
]

@@

---
# Rural Shrink Smart Project
.center[
![](Shrink_Smart_logo.png)
]
.center[
(https://ruralshrinksmart.org/)
]

@@

.medium.emph[
This material is based upon work supported by the National Science Foundation under Grant No. 1952007. We would also like to thank our collaborators at Iowa State, the Iowa League of Cities, and the various Iowa government organizations which publish their data at (https://data.iowa.gov).
]

---
# Complications

@@

---
# Complications

@@

---
# Basic Idea

@@

- Assess usability, interpretability and adaptability to new statistical methodology.

- .emph.cerulean[The Rural Shrink Smart EDA dashboard incorporates two major components]: 
    - Data Visualization Interface component with Guided Discovery Learning (GDL) principles
    - Exploratory Data Analysis component utilizing statistical clustering methods.

---
#

---
# Guiding Design Principles

Our initial set of dashboard design principles are:

- The town leaders are the focus audience (central focus of the app)

- Facilitate comparisons with other towns in order to allow the user to explore other potential solutions

- Present the user with peer comparisons in order to widen the scope of exploration beyond the initial set of obvious peers in the local region.

- Allow for more detailed data and feature requests to improve the dashboard design over time.

---
# Guided Discovery Learning with Dashboard Design

We leverage the framework of Guided Discovery Learning (GDL) to guide the town leaders to make discoveries using our interactive visualization. By leveraging:

- hints

- feedback

- other helpful information to guide users in interactive exploration
    e.g. user ignoring the radio buttons to compare similar cities, etc.

---
# Dashboard Design Workflow

@@

- Dashboard layout
    - Which layout will allow the user to engage the most without overwhelming their curiosity?

- Dashboard Feedback
    - What elements on the dashboard are useful, confusing, misleading or incorrect from town leaders?

---
# Visualization

@@

---
# Dashboard Visual Iterations

- Version 1: An idea without limitations of the actual implementation pieces.

- Version 2: After implementation has been completed and the team notes limitations, we change the interface.

- Version 3: Now incorporate more literature on dashboard design and user experience, which removes some of the "cool" implementations

- Version 4: Finally incorporate a collective feedback from team-leaders that suggests lack of understanding of information

#
